Alex's Notes and Text Fragments for the JSS Manuscript

### 2021-08-04 paper/paper.Rmd
revised coordinate-based clustering section

### Clustering

Cluster analysis groups objects with similar characteristics.
In the spatial context, the similarity of spatial coordinates is used to form such groups.
Different cluster algorithms such as hierarchical, density-based, grid-based or centroid-based clustering exist [@gower1967].

#### Coordinate-based clustering --- sperrorest

Coordinate-based clustering for spatial CV [@russ2010; @brenning2012] as implemented in package \pkg{sperrorest} uses the coordinates of all observations to create clusters in the spatial domain with the help of the $k$-means clustering algorithm.
To avoid confusion with other spatial partitioning methods, we refer to this as coordinate-based clustering, or as `spcv_coords` in \pkg{mlr3spatiotempcv}, instead of using the generic "Spatial CV" term.

This approach is very versatile as it adapts to irregularly-shaped study areas and ensures that exactly $k$ partitions are created.
Nevertheless, despite the random selection of initial cluster centers, repeated partitionings may in some cases be nearly identical.
Also, $k$-means clustering may be less suitable for data sets with existing clusters of points and/or with distant isolated sample locations.

@@

### Custom and grouped resampling --- mlr3

>>>>

### Clustering-based resampling

Cluster analysis provides a flexible approach to creating irregularly shaped spatial blocks for spatial resampling.
Numerous techniques are available that can potentially be applied to the spatial coordinates of observations, to the features, or to a combination of both.
In spatial model assessment, the focus has been on coordinate-based clustering, and specifically on leave-one-block-out resampling with blocks created by $k$-means clustering of the coordinates [@russ2010].

#### Leave-one-block-out resampling using coordinate-based clustering --- sperrorest

Coordinate-based clustering for spatial CV [@russ2010; @brenning2012] as implemented in package \pkg{sperrorest} uses the coordinates of all observations to create clusters in the spatial domain with the help of the $k$-means clustering algorithm.
This can be regarded as a leave-one-block-out resampling method, or as a $k$-fold CV in which each test set is a spatial cluster. 
This method is referred to as `spcv_coords` in \pkg{mlr3spatiotempcv}.

Coordinate-based clustering approach is very versatile as it adapts to irregularly-shaped study areas and ensures that exactly $k$ partitions are created, which are usually or very similar size when the sample locations are spread out evenly.
Nevertheless, despite the random selection of initial cluster centers, repeated partitionings may in some cases be nearly identical.
Also, $k$-means clustering may be less suitable for data sets with pre-existing clusters of points and/or with distant isolated sample locations.
When distinct clusters of points are present, as in multi-level sampling, it may be better to define clusters using a factor variable (see method `"custom_cv"` in Section \@ref(sec:custom-cv)).

@@

### Custom and grouped resampling --- mlr3 {#sec:custom-cv}
